Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2024-09-22 article.tex
Updates for Sec 4

The equivalent (von Mises) stress at full inflation is shown for the mesh with 829,440 cells in Figure \ref{fig:ventricle}(b), where the high stresses on the endocardial (inner) surface quickly drop off through the wall thickness towards the endocardial (outer) surface.
The predicted deformed configuration of the midline of the ventricle wall is shown in Figure \ref{fig:ventricle_accuracy}, where a side-by-side comparison is given with the round-robin benchmark results from \citet{Land2015}.
The predictions are seen to become quickly mesh-independent and fall within the benchmark ranges.
%The variation in the round-robin results may be explained by the sensitivity of the predictions to the allowed compressibility; when a penalty bulk modulus approach is adopted, like in the current work, increasing
\begin{figure}[htbp]
	\centering
	\subfigure[Predictions from the current work]
	{
   		\includegraphics[width=0.46\textwidth]{figures/ventricle_results} 
   	}
	\subfigure[\citet{Land2015} benchmarks]
	{
   		\includegraphics[width=0.48\textwidth]{figures/ventricle_results_land2015}  
   	}
	\caption{Idealised ventricle: predictions for the current method (left) vs the round-robin predictions from \citet{Land2015} \hl{Merge figures}}
	\label{fig:ventricle_accuracy}
\end{figure}

\paragraph{Case 5: Cook's membrane}
Cook's membrane (Figure \ref{fig:cooks_membrane}) is a well-known bending-dominated benchmark case used in linear and non-linear analysis.
The current work considers three forms of the problem: (i) small strain elastic, (ii) finite-strain hyperelastic, and (iii) finite strain elastoplastic version \citep{Simo1992}.
The 2-D plane strain tapered panel (trapezoid) is fixed on one side and subjected to uniform shear traction on the opposite side.
The prescribed shear traction is $\tau = 0.3125$ MPa.
The vertices of the trapezoid (in mm) are (0, 0), (48, 44), (48, 60),  and (0, 44).
The problem is solved as static, using 30 equally-sized loading increments, and there are no body forces.
%\begin{figure}[htbp]
%   \centering
%   \includegraphics[width=0.4\textwidth]{figures/cooksMembrane-geometry} 
%   \caption{Cook's membrane case geometry (dimensions in mm) and loading conditions (fixed left, traction $\tau$ applied to right)}
%   \label{fig:cooks_membrane}
%\end{figure}
\begin{figure}[H]
	\centering
	\subfigure[Case geometry and dimensions]
	{
		\label{fig:cooks_membrane_geometry}
   		\includegraphics[scale=1]{figures/cooksMembrane-geometry} 
   	}
   	\qquad
	\subfigure[Hexahedral mesh with 144 cells]
	{
		\label{fig:cooks_membrane_mesh}
   		\includegraphics[scale=0.14]{figures/cooksMembrane-mesh}  
   	}
	\caption{Cook's membrane case geometry and mesh}
   \label{fig:cooks_membrane}
\end{figure}

\hl{Define properties for three cases: i, ii and iii}
The Young's modulus $E = 206.9$ MPa and Poisson's ratio $\nu=0.29$, with the yield stress $\sigma_y$ given as \citep{Simo1992}
\begin{eqnarray}
	\sigma_y = \sigma_Y + H\bar{\varepsilon}_p + (\sigma_{\infty} - \sigma_Y)(1- e^{-\delta\bar{\varepsilon}_p})
\end{eqnarray}
The plastic yielding parameters are $\sigma_Y = 0.45$ MPa, $\sigma_{\infty} = 0.715$ MPa, $\delta = 16.93$, and $H = 0.12924$ MPa, where the hardening variable $\bar{\varepsilon}_p$ corresponds to equivalent plastic strain.

\hl{Give mesh details}

\hl{Add figures: finest mesh stress distributions and deformation}
\hl{Add figures: corner displacement prediction vs average cell width}

>>>>

The equivalent (von Mises) stress at full inflation is shown for the mesh with 829,440 cells in Figure \ref{fig:ventricle}(b), where the high stresses on the endocardial (inner) surface quickly drop off through the wall thickness towards the endocardial (outer) surface.
The predicted deformed configuration of the midline of the ventricle wall is shown in Figure \ref{fig:ventricle_accuracy}, where a side-by-side comparison is given with the round-robin benchmark results from \citet{Land2015}.
The predictions are seen to become quickly mesh-independent and fall within the benchmark ranges.
%The variation in the round-robin results may be explained by the sensitivity of the predictions to the allowed compressibility; when a penalty bulk modulus approach is adopted, like in the current work, increasing
\begin{figure}[htbp]
	\centering
   		\includegraphics[width=0.6\textwidth]{figures/ventricle_results} 
%	\subfigure[Predictions from the current work]
%	{
%   		\includegraphics[width=0.46\textwidth]{figures/ventricle_results} 
%   	}
%	\subfigure[\citet{Land2015} benchmarks]
%	{
%		\includegraphics[width=0.48\textwidth]{figures/ventricle_results_land2015}  
%	}
	\caption{Idealised ventricle: predictions for the four meshes from the present work overlaid on the round-robin predictions from \citet{Land2015}}
	\label{fig:ventricle_accuracy}
\end{figure}

\paragraph{Case 5: Cook's membrane}
Cook's membrane (Figure \ref{fig:cooks_membrane}) is a well-known bending-dominated benchmark case used in linear and non-linear analysis.
The 2-D plane strain tapered panel (trapezoid) is fixed on one side and subjected to uniform shear traction on the opposite side.
The vertices of the trapezoid (in mm) are (0, 0), (48, 44), (48, 60),  and (0, 44).
The problem is solved quasi-statically, using 30 equally-sized loading increments, and there are no body forces.
%\begin{figure}[htbp]
%   \centering
%   \includegraphics[width=0.4\textwidth]{figures/cooksMembrane-geometry} 
%   \caption{Cook's membrane case geometry (dimensions in mm) and loading conditions (fixed left, traction $\tau$ applied to right)}
%   \label{fig:cooks_membrane}
%\end{figure}
\begin{figure}[htbp]
	\centering
	\subfigure[Case geometry and dimensions]
	{
		\label{fig:cooks_membrane_geometry}
   		\includegraphics[scale=1]{figures/cooksMembrane-geometry} 
   	}
   	\qquad
	\subfigure[Hexahedral mesh with 144 cells]
	{
		\label{fig:cooks_membrane_mesh}
   		\includegraphics[scale=0.14]{figures/cooksMembrane-mesh}  
   	}
	\caption{Cook's membrane case geometry and mesh}
   \label{fig:cooks_membrane}
\end{figure}

The current work considers three forms of the problem:
\begin{enumerate}[label=\roman*.]
	\item Small strain linear elastic \cite{Zienkiewicz2000}: $E=70$ MPa, $\nu=1/3$, and $\tau = 6.25$ MPa.
	\item Finite strain neo-Hookean hyperelastic \cite{dealII95}: $E=1.0985$ MPa, $\nu=0.3$, and $\tau = 0.0625$ MPa.
	\item Finite strain elastoplastic \citep{Simo1992}: $E=206.9$ MPa, $\nu=0.29$, $\sigma_y = 0.45 + 0.12924\bar{\varepsilon}_p + (0.715 - 0.45)(1- e^{-16.93\bar{\varepsilon}_p})$ MPa, and $\tau = 0.3125$ MPa.
\end{enumerate}
where $E$ is the Young's modulus, $\nu$ is the Poisson's ratio, $\sigma_y$ is the yield strength, $\bar{\varepsilon}_p$ is the equivalent plastic strain, and $\tau$ is the prescribed shear traction.

%The Young's modulus $E = 206.9$ MPa and Poisson's ratio $\nu=0.29$, with the yield stress $\sigma_y$ given as \citep{Simo1992}
%\begin{eqnarray}
%	\sigma_y = \sigma_Y + H\bar{\varepsilon}_p + (\sigma_{\infty} - \sigma_Y)(1- e^{-\delta\bar{\varepsilon}_p})
%\end{eqnarray}
%The plastic yielding parameters are $\sigma_Y = 0.45$ MPa, $\sigma_{\infty} = 0.715$ MPa, $\delta = 16.93$, and $H = 0.12924$ MPa, where the hardening variable $\bar{\varepsilon}_p$ corresponds to equivalent plastic strain.
%\noindent
%\textbf{Sources for case i:} \cite{Zienkiewicz2000}\\
%O.C. Zienkiewicz, R.L. Taylor. The finite element method. Butterworth Heinemann, 2000.\\
%https://www.simplassoftware.com/benchmarks.html\#biblio-58\\
%https://github.com/spolanski/CoFEA/tree/master/benchmarks/02-cooks-membrane\\
%\textbf{Sources for case ii:} \cite{dealII95}\\
%https://www.dealii.org/current/doxygen/deal.II/code\_gallery\_Quasi\_static\_Finite\_strain\_Compressible\_Elasticity.html\\
%\textbf{Sources for case iii:} \citep{Simo1992}

Seven successively refined quadrilateral meshes are considered, where the cell counts are 9, 36, 144 (Figure \ref{fig:cooks_membrane}(b)), 576, $2\,304$, $9\,216$, and $36\,864$.
%mesh.1 - 3x3 - 9 CVs\\
%mesh.2 - 6x6 - 36 CVs\\
%mesh.3 - 12x12 - 144 CVs\\
%mesh.4 - 24x24 - 576 CVs\\
%mesh.5 - 48-48 - 2304 CVs\\
%mesh.6 - 96x96 - 9216 CVs\\
%mesh.7 - 192x192 - 36864 CVs

%\hl{Add figures: finest mesh stress distributions and deformation Add figures: corner displacement prediction vs average cell width}
\hl{For this section, I think it is fine if we only discuss the hex.snes results (and maybe hex.snes.seg)}
\hl{If hex.snes diverges for case iii, we can state that and give on the preconditioner}
\hl{I suggest we show sigmaEq for all three cases => case (iii) may have to be hex.snes.seg results}

Figure \ref{fig:cooksMembrane_sigmaEq} shows the predicted equivalent stress distribution on the mesh with $36\,842$ cells for cases i, ii and iii.
\hl{Add a comment on the distributions}

\begin{figure}[htbp]
   \centering
	\subfigure[Linear elastic case]
	{
		\includegraphics[width=0.3\textwidth]{figures/placeholder}  
	}
	\subfigure[Hyperelastic case]
	{
		\includegraphics[width=0.3\textwidth]{figures/cooksMembrane-sigmaEq.pdf}  
	}
	\subfigure[Hyperelastoplastic case]
	{
		\includegraphics[width=0.3\textwidth]{figures/placeholder}  
	}
   \caption{Equivalent (von Mises) stress distribution for the three cases using the mesh with $36\,842$ cells}
   \label{fig:cooksMembrane_sigmaEq}
\end{figure}

\hl{I suggest we show a comparison to literature to show we get the same answers as others}
Figure \ref{fig:cooksMembrane_disp} presents the predicted vertical displacement for the top right point (48,60) vs the average cell widths, where results from the literature are given for comparison.
\hl{Comment on results -> good/bad/etc}
\hl{Comment on JFNK not working for plasticty}
\begin{figure}[htbp]
   \centering
		\includegraphics[width=0.5\textwidth]{figures/cooksMembrane-tipDispConvergence.pdf}  
%	\subfigure[Equivalent stress distribution for the hyperelastic case (case ii) using the mesh with $36\,842$ cells]
%	{
%		\includegraphics[height=0.5\textwidth]{figures/cooksMembrane-sigmaEq.pdf}  
%	}
%	\subfigure[Convergence of tip (48,60) vertical displacement, normalised using the displacement value from the finest mesh (given in parentheses)]
%	{
%		\includegraphics[height=0.35\textwidth]{figures/cooksMembrane-tipDispConvergence.pdf} 
%	}
   \caption{Cook's membrane top right corner vertical displacement predictions vs the average cell width. Comparisons are given with the results from \hl{XX} and \hl{YY}}
   \label{fig:cooksMembrane_disp}
\end{figure}
